---
title: "251206_figures_analyses2"
output: html_document
date: "2025-12-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary libraries
add additional packages as necessary when additional packages are needed
```{r}
library(survminer)
library(survivalAnalysis)
library(magrittr)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(readxl)
library(plyr)
library(cmprsk)
library(cowplot)
library(ggplot2)
library(AICcmodavg)
library(cmprsk)
```

Load the dataframe
Note: for simplicity reason we are naming the full dataframe "df" 
```{r}
df<- read_xlsx("/Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/MRDMRGMshared/Export data for R/251110_MRD_MRG_t2tx_v2_anonym.xlsx", na = "na")
```

Transform all relevant variables to numbers
```{r}
df$`Einschluss t1_oder_t2`<-as.numeric(df$`Einschluss t1_oder_t2`)
df$Tx_yes_no<-as.numeric(df$Tx_yes_no)
df$t1_MRD_Comutations_positive1_negative0<-as.numeric(df$t1_MRD_Comutations_positive1_negative0)
df$t1_MRD_MRGM_positive1_negative0<-as.numeric(df$t1_MRD_MRGM_positive1_negative0)
df$t2_MRD_Comutations_positive1_negative0<-as.numeric(df$t2_MRD_Comutations_positive1_negative0)
df$t2_MRD_MRGM_positive1_negative0<-as.numeric(df$t2_MRD_MRGM_positive1_negative0)
df$ELN<-as.numeric(df$ELN)
df$`CBF, NPM1, TP53, adverse`<-as.numeric(df$`CBF, NPM1, TP53, adverse`)
df$Age_at_ED<-as.numeric(df$Age_at_ED)
df$Sex<-as.numeric(df$Sex)
df$NPM1_code<-as.numeric(df$NPM1_code)
df$TP53_code<-as.numeric(df$TP53_code)
df$NPM1_code<-as.numeric(df$NPM1_code)
df$`t1_eingeschlossen 1 _ ausgeschlossen 0`<-as.numeric(df$`t1_eingeschlossen 1 _ ausgeschlossen 0`)
df$t2_eingeschlossen1_ausgeschlossen0<-as.numeric(df$t2_eingeschlossen1_ausgeschlossen0)
df$ECOG_Code<-as.numeric(df$ECOG_Code)
df$WBC_at_ED<-as.numeric(df$WBC_at_ED)
df$AML_Type<-as.numeric(df$AML_Type)
df$KT_complex<-as.numeric(df$KT_complex)
df$Hb_at_ED<-as.numeric(df$Hb_at_ED)
df$Platelets_at_ED<-as.numeric(df$Platelets_at_ED)
df$Type_of_consolidation_therapy<-as.numeric(df$Type_of_consolidation_therapy)
df$t1_MRD_MRGM_noASXL1<-as.numeric(df$t1_MRD_MRGM_noASXL1)
df$t2_MRD_MRGM_noASXL1<-as.numeric(df$t2_MRD_MRGM_noASXL1)
df$t1_FLT3_MRD<-as.numeric(df$t1_FLT3_MRD)
df$t2_FLT3_MRD<-as.numeric(df$t2_FLT3_MRD)
df$t1_NPM1_MRD<-as.numeric(df$t1_NPM1_MRD)
df$t2_NPM1_MRD<-as.numeric(df$t2_NPM1_MRD)
df$CIR_event<-as.numeric(df$CIR_event)
df$t1_MRD_ASXL1<-as.numeric(df$t1_MRD_ASXL1)
df$t1_MRD_VAF_ASXL1<-as.numeric(df$t1_MRD_VAF_ASXL1)
df$t1_MRD_BCOR<-as.numeric(df$t1_MRD_BCOR)
df$t1_MRD_VAF_BCOR<-as.numeric(df$t1_MRD_VAF_BCOR)
df$t1_MRD_EZH2<-as.numeric(df$t1_MRD_EZH2)
df$t1_MRD_VAF_EZH2<-as.numeric(df$t1_MRD_VAF_EZH2)
df$t1_MRD_RUNX1<-as.numeric(df$t1_MRD_RUNX1)
df$t1_MRD_VAF_RUNX1<-as.numeric(df$t1_MRD_VAF_RUNX1)
df$t1_MRD_SF3B1<-as.numeric(df$t1_MRD_SF3B1)
df$t1_MRD_VAF_SF3B1<-as.numeric(df$t1_MRD_VAF_SF3B1)
df$t1_MRD_SRSF2<-as.numeric(df$t1_MRD_SRSF2)
df$t1_MRD_VAF_SRSF2<-as.numeric(df$t1_MRD_VAF_SRSF2)
df$t1_MRD_STAG2<-as.numeric(df$t1_MRD_STAG2)
df$t1_MRD_VAF_STAG2<-as.numeric(df$t1_MRD_VAF_STAG2)
df$t1_MRD_U2AF1<-as.numeric(df$t1_MRD_U2AF1)
df$t1_MRD_VAF_U2AF1<-as.numeric(df$t1_MRD_VAF_U2AF1)
df$t1_MRD_ZRSR2<-as.numeric(df$t1_MRD_ZRSR2)
df$t1_MRD_VAF_ZRSR2<-as.numeric(df$t1_MRD_VAF_ZRSR2)
df$t2_MRD_ASXL1<-as.numeric(df$t2_MRD_ASXL1)
df$t2_MRD_VAF_ASXL1<-as.numeric(df$t2_MRD_VAF_ASXL1)
df$t2_MRD_BCOR<-as.numeric(df$t2_MRD_BCOR)
df$t2_MRD_VAF_BCOR<-as.numeric(df$t2_MRD_VAF_BCOR)
df$t2_MRD_EZH2<-as.numeric(df$t2_MRD_EZH2)
df$t2_MRD_VAF_EZH2<-as.numeric(df$t2_MRD_VAF_EZH2)
df$t2_MRD_RUNX1<-as.numeric(df$t2_MRD_RUNX1)
df$t2_MRD_VAF_RUNX1<-as.numeric(df$t2_MRD_VAF_RUNX1)
df$t2_MRD_SF3B1<-as.numeric(df$t2_MRD_SF3B1)
df$t2_MRD_VAF_SF3B1<-as.numeric(df$t2_MRD_VAF_SF3B1)
df$t2_MRD_SRSF2<-as.numeric(df$t2_MRD_SRSF2)
df$t2_MRD_VAF_SRSF2<-as.numeric(df$t2_MRD_VAF_SRSF2)
df$t2_MRD_STAG2<-as.numeric(df$t2_MRD_STAG2)
df$t2_MRD_VAF_STAG2<-as.numeric(df$t2_MRD_VAF_STAG2)
df$t2_MRD_U2AF1<-as.numeric(df$t2_MRD_U2AF1)
df$t2_MRD_VAF_U2AF1<-as.numeric(df$t2_MRD_VAF_U2AF1)
df$t2_MRD_ZRSR2<-as.numeric(df$t2_MRD_ZRSR2)
df$t2_MRD_VAF_ZRSR2<-as.numeric(df$t2_MRD_VAF_ZRSR2)
df$t1_MRD_Comutations_noDTA<-as.numeric(df$t1_MRD_Comutations_noDTA)
df$t2_MRD_Comutations_noDTA<-as.numeric(df$t2_MRD_Comutations_noDTA)
df$`DTA?`<-as.numeric(df$`DTA?`)
df$Sample_Material_PB_1_BM_2<-as.numeric(df$Sample_Material_PB_1_BM_2)
df$Sample_Material_t2_PB_1_BM_2<-as.numeric(df$Sample_Material_t2_PB_1_BM_2)
df$t1_MRD_ATRX <- as.numeric(df$t1_MRD_ATRX)
df$t1_MRD_VAF_ATRX <- as.numeric(df$t1_MRD_VAF_ATRX)
df$t1_MRD_BCORL1 <- as.numeric(df$t1_MRD_BCORL1)
df$t1_MRD_VAF_BCORL1 <- as.numeric(df$t1_MRD_VAF_BCORL1)
df$t1_MRD_CALR <- as.numeric(df$t1_MRD_CALR)
df$t1_MRD_VAF_CALR <- as.numeric(df$t1_MRD_VAF_CALR)
df$t1_MRD_CBL <- as.numeric(df$t1_MRD_CBL)
df$t1_MRD_VAF_CBL <- as.numeric(df$t1_MRD_VAF_CBL)
df$t1_MRD_CEBPA <- as.numeric(df$t1_MRD_CEBPA)
df$t1_MRD_VAF_CEBPA <- as.numeric(df$t1_MRD_VAF_CEBPA)
df$t1_MRD_CSF3R <- as.numeric(df$t1_MRD_CSF3R)
df$t1_MRD_VAF_CSF3R <- as.numeric(df$t1_MRD_VAF_CSF3R)
df$t1_MRD_DDX41 <- as.numeric(df$t1_MRD_DDX41)
df$t1_MRD_VAF_DDX41 <- as.numeric(df$t1_MRD_VAF_DDX41)
df$t1_MRD_DNMT3A <- as.numeric(df$t1_MRD_DNMT3A)
df$t1_MRD_VAF_DNMT3A <- as.numeric(df$t1_MRD_VAF_DNMT3A)
df$t1_MRD_ETNK1 <- as.numeric(df$t1_MRD_ETNK1)
df$t1_MRD_VAF_ETNK1 <- as.numeric(df$t1_MRD_VAF_ETNK1)
df$t1_MRD_ETV6 <- as.numeric(df$t1_MRD_ETV6)
df$t1_MRD_VAF_ETV6 <- as.numeric(df$t1_MRD_VAF_ETV6)
df$t1_MRD_FLT3 <- as.numeric(df$t1_MRD_FLT3)
df$t1_MRD_VAF_FLT3 <- as.numeric(df$t1_MRD_VAF_FLT3)
df$t1_MRD_GATA2 <- as.numeric(df$t1_MRD_GATA2)
df$t1_MRD_VAF_GATA2 <- as.numeric(df$t1_MRD_VAF_GATA2)
df$t1_MRD_IDH1 <- as.numeric(df$t1_MRD_IDH1)
df$t1_MRD_VAF_IDH1 <- as.numeric(df$t1_MRD_VAF_IDH1)
df$t1_MRD_IDH2 <- as.numeric(df$t1_MRD_IDH2)
df$t1_MRD_VAF_IDH2 <- as.numeric(df$t1_MRD_VAF_IDH2)
df$t1_MRD_JAK2 <- as.numeric(df$t1_MRD_JAK2)
df$t1_MRD_VAF_JAK2 <- as.numeric(df$t1_MRD_VAF_JAK2)
df$t1_MRD_KDM6A <- as.numeric(df$t1_MRD_KDM6A)
df$t1_MRD_VAF_KDM6A <- as.numeric(df$t1_MRD_VAF_KDM6A)
df$t1_MRD_KIT <- as.numeric(df$t1_MRD_KIT)
df$t1_MRD_VAF_KIT <- as.numeric(df$t1_MRD_VAF_KIT)
df$t1_MRD_KMT2A <- as.numeric(df$t1_MRD_KMT2A)
df$t1_MRD_VAF_KMT2A <- as.numeric(df$t1_MRD_VAF_KMT2A)
df$t1_MRD_KRAS <- as.numeric(df$t1_MRD_KRAS)
df$t1_MRD_VAF_KRAS <- as.numeric(df$t1_MRD_VAF_KRAS)
df$t1_MRD_MYC <- as.numeric(df$t1_MRD_MYC)
df$t1_MRD_VAF_MYC <- as.numeric(df$t1_MRD_VAF_MYC)
df$t1_MRD_NF1 <- as.numeric(df$t1_MRD_NF1)
df$t1_MRD_VAF_NF1 <- as.numeric(df$t1_MRD_VAF_NF1)
df$t1_MRD_NPM1 <- as.numeric(df$t1_MRD_NPM1)
df$t1_MRD_VAF_NPM1 <- as.numeric(df$t1_MRD_VAF_NPM1)
df$t1_MRD_NRAS <- as.numeric(df$t1_MRD_NRAS)
df$t1_MRD_VAF_NRAS <- as.numeric(df$t1_MRD_VAF_NRAS)
df$t1_MRD_PHF6 <- as.numeric(df$t1_MRD_PHF6)
df$t1_MRD_VAF_PHF6 <- as.numeric(df$t1_MRD_VAF_PHF6)
df$t1_MRD_PTPN11 <- as.numeric(df$t1_MRD_PTPN11)
df$t1_MRD_VAF_PTPN11 <- as.numeric(df$t1_MRD_VAF_PTPN11)
df$t1_MRD_RAD21 <- as.numeric(df$t1_MRD_RAD21)
df$t1_MRD_VAF_RAD21 <- as.numeric(df$t1_MRD_VAF_RAD21)
df$t1_MRD_SETBP1 <- as.numeric(df$t1_MRD_SETBP1)
df$t1_MRD_VAF_SETBP1 <- as.numeric(df$t1_MRD_VAF_SETBP1)
df$t1_MRD_SMC1A <- as.numeric(df$t1_MRD_SMC1A)
df$t1_MRD_VAF_SMC1A <- as.numeric(df$t1_MRD_VAF_SMC1A)
df$t1_MRD_SMC3 <- as.numeric(df$t1_MRD_SMC3)
df$t1_MRD_VAF_SMC3 <- as.numeric(df$t1_MRD_VAF_SMC3)
df$t1_MRD_STAG1 <- as.numeric(df$t1_MRD_STAG1)
df$t1_MRD_VAF_STAG1 <- as.numeric(df$t1_MRD_VAF_STAG1)
df$t1_MRD_TET2 <- as.numeric(df$t1_MRD_TET2)
df$t1_MRD_VAF_TET2 <- as.numeric(df$t1_MRD_VAF_TET2)
df$t1_MRD_WT1 <- as.numeric(df$t1_MRD_WT1)
df$t1_MRD_VAF_WT1 <- as.numeric(df$t1_MRD_VAF_WT1)
df$t2_MRD_ATRX <- as.numeric(df$t2_MRD_ATRX)
df$t2_MRD_VAF_ATRX <- as.numeric(df$t2_MRD_VAF_ATRX)
df$t2_MRD_BCORL1 <- as.numeric(df$t2_MRD_BCORL1)
df$t2_MRD_VAF_BCORL1 <- as.numeric(df$t2_MRD_VAF_BCORL1)
df$t2_MRD_CALR <- as.numeric(df$t2_MRD_CALR)
df$t2_MRD_VAF_CALR <- as.numeric(df$t2_MRD_VAF_CALR)
df$t2_MRD_CBL <- as.numeric(df$t2_MRD_CBL)
df$t2_MRD_VAF_CBL <- as.numeric(df$t2_MRD_VAF_CBL)
df$t2_MRD_CEBPA <- as.numeric(df$t2_MRD_CEBPA)
df$t2_MRD_VAF_CEBPA <- as.numeric(df$t2_MRD_VAF_CEBPA)
df$t2_MRD_CSF3R <- as.numeric(df$t2_MRD_CSF3R)
df$t2_MRD_VAF_CSF3R <- as.numeric(df$t2_MRD_VAF_CSF3R)
df$t2_MRD_DDX41 <- as.numeric(df$t2_MRD_DDX41)
df$t2_MRD_VAF_DDX41 <- as.numeric(df$t2_MRD_VAF_DDX41)
df$t2_MRD_DNMT3A <- as.numeric(df$t2_MRD_DNMT3A)
df$t2_MRD_VAF_DNMT3A <- as.numeric(df$t2_MRD_VAF_DNMT3A)
df$t2_MRD_ETNK1 <- as.numeric(df$t2_MRD_ETNK1)
df$t2_MRD_VAF_ETNK1 <- as.numeric(df$t2_MRD_VAF_ETNK1)
df$t2_MRD_FLT3 <- as.numeric(df$t2_MRD_FLT3)
df$t2_MRD_VAF_FLT3 <- as.numeric(df$t2_MRD_VAF_FLT3)
df$t2_MRD_GATA2 <- as.numeric(df$t2_MRD_GATA2)
df$t2_MRD_VAF_GATA2 <- as.numeric(df$t2_MRD_VAF_GATA2)
df$t2_MRD_IDH1 <- as.numeric(df$t2_MRD_IDH1)
df$t2_MRD_VAF_IDH1 <- as.numeric(df$t2_MRD_VAF_IDH1)
df$t2_MRD_IDH2 <- as.numeric(df$t2_MRD_IDH2)
df$t2_MRD_VAF_IDH2 <- as.numeric(df$t2_MRD_VAF_IDH2)
df$t2_MRD_JAK2 <- as.numeric(df$t2_MRD_JAK2)
df$t2_MRD_VAF_JAK2 <- as.numeric(df$t2_MRD_VAF_JAK2)
df$t2_MRD_MYC <- as.numeric(df$t2_MRD_MYC)
df$t2_MRD_VAF_MYC <- as.numeric(df$t2_MRD_VAF_MYC)
df$t2_MRD_NF1 <- as.numeric(df$t2_MRD_NF1)
df$t2_MRD_VAF_NF1 <- as.numeric(df$t2_MRD_VAF_NF1)
df$t2_MRD_NPM1 <- as.numeric(df$t2_MRD_NPM1)
df$t2_MRD_VAF_NPM1 <- as.numeric(df$t2_MRD_VAF_NPM1)
df$t2_MRD_NRAS <- as.numeric(df$t2_MRD_NRAS)
df$t2_MRD_VAF_NRAS <- as.numeric(df$t2_MRD_VAF_NRAS)
df$t2_MRD_PHF6 <- as.numeric(df$t2_MRD_PHF6)
df$t2_MRD_VAF_PHF6 <- as.numeric(df$t2_MRD_VAF_PHF6)
df$t2_MRD_RAD21 <- as.numeric(df$t2_MRD_RAD21)
df$t2_MRD_VAF_RAD21 <- as.numeric(df$t2_MRD_VAF_RAD21)
df$t2_MRD_SMC3 <- as.numeric(df$t2_MRD_SMC3)
df$t2_MRD_VAF_SMC3 <- as.numeric(df$t2_MRD_VAF_SMC3)
df$t2_MRD_TET2 <- as.numeric(df$t2_MRD_TET2)
df$t2_MRD_VAF_TET2 <- as.numeric(df$t2_MRD_VAF_TET2)
df$t2_MRD_WT1 <- as.numeric(df$t2_MRD_WT1)
df$t2_MRD_VAF_WT1 <- as.numeric(df$t2_MRD_VAF_WT1)
df$numberofdays_tx_t2sample <- as.numeric(df$numberofdays_tx_t2sample)
df$tx_after_relapse <- as.numeric(df$tx_after_relapse)
df$RFS_years_censored <- as.numeric(df$RFS_years_censored)
df$RFS_MRDt2_years <- as.numeric(df$RFS_MRDt2_years)
df$RFS_MRDt2_years_censored <- as.numeric(df$RFS_MRDt2_years_censored)
df$RFS_MRDt1_years <- as.numeric(df$RFS_MRDt1_years)
df$RFS_MRDt1_years_censored <- as.numeric(df$RFS_MRDt1_years_censored)
df$OS_years_censored <- as.numeric(df$OS_years_censored)
df$OS_MRDt1_days <- as.numeric(df$OS_MRDt1_days)
df$OS_MRDt1_years <- as.numeric(df$OS_MRDt1_years)
df$OS_MRDt1_years_censored <- as.numeric(df$OS_MRDt1_years_censored)
df$OS_MRDt2_days <- as.numeric(df$OS_MRDt2_days)
df$OS_MRDt2_years <- as.numeric(df$OS_MRDt2_years)
df$OS_MRDt2_years_censored <- as.numeric(df$OS_MRDt2_years_censored)
df$Txin_CR1_vs_CR2 <- as.numeric(df$Txin_CR1_vs_CR2)
df$OS_fromTx_years <- as.numeric(df$OS_fromTx_years)
df$OS_fromTx_years_censored <- as.numeric(df$OS_fromTx_years_censored)
df$RFS_fromTx_years <- as.numeric(df$RFS_fromTx_years)
df$RFS_fromTx_years_censored <- as.numeric(df$RFS_fromTx_years_censored)
df$MRD_conversion <- as.numeric(df$MRD_conversion)
df$Tx_Salvage <- as.numeric(df$Tx_Salvage)
df$Tx_Salvage_Dat <- as.numeric(df$Tx_Salvage_Dat)
df$best_response_salvage <- as.numeric(df$best_response_salvage)
df$EDtoTx_indays <- as.numeric(df$EDtoTx_indays)
df$Ind1toTx_indays <- as.numeric(df$Ind1toTx_indays)
df$Ind2toTx_indays <- as.numeric(df$Ind2toTx_indays)
df$Cons1toTx_indays <- as.numeric(df$Cons1toTx_indays)
df$Pat_Age_atTx1 <- as.numeric(df$Pat_Age_atTx1)
df$Patient_Sex_atTx1 <- as.numeric(df$Patient_Sex_atTx1)
df$Donor_Sex_atTx1 <- as.numeric(df$Donor_Sex_atTx1)
df$Donor_match_Tx1 <- as.numeric(df$Donor_match_Tx1)
df$Donor_match_related_vs_unrelated <- as.numeric(df$Donor_match_related_vs_unrelated)
df$CMVstatus_Pat_atTx1 <- as.numeric(df$CMVstatus_Pat_atTx1)
df$CMVstatus_Don_atTx1 <- as.numeric(df$CMVstatus_Don_atTx1)
df$CMVstatus_PnegDneg_vs_rest <- as.numeric(df$CMVstatus_PnegDneg_vs_rest)
df$CMV_4groups_code <- as.numeric(df$CMV_4groups_code)
df$StemCell_Source <- as.numeric(df$StemCell_Source)
df$Blasts_before_Tx1_PB <- as.numeric(df$Blasts_before_Tx1_PB)
df$WBC_before_Tx1 <- as.numeric(df$WBC_before_Tx1)
df$Hb_before_Tx1 <- as.numeric(df$Hb_before_Tx1)
df$Platelets_before_Tx1 <- as.numeric(df$Platelets_before_Tx1)
df$GvHD_allg <- as.numeric(df$GvHD_allg)
df$GvHD_afterDLI <- as.numeric(df$GvHD_afterDLI)
df$aGvHD_Tx1 <- as.numeric(df$aGvHD_Tx1)
df$aGvHD_Tx2 <- as.numeric(df$aGvHD_Tx2)
df$aGvHDmax_grade <- as.numeric(df$aGvHDmax_grade)
df$aGvHD_grade_low_vs_high <- as.numeric(df$aGvHD_grade_low_vs_high)
df$cGvHD_Tx1 <- as.numeric(df$cGvHD_Tx1)
df$cGvHD_Tx2 <- as.numeric(df$cGvHD_Tx2)
df$cGvHD_Tx1_grade <- as.numeric(df$cGvHD_Tx1_grade)
df$cycles_before_Tx_1_2_3ormore <- as.numeric(df$cycles_before_Tx_1_2_3ormore)
df$HCTCI_at_ED <- as.numeric(df$HCTCI_at_ED)
df$StemCell_Source <- as.numeric(df$StemCell_Source)
df$`ICC classifier` <- as.numeric(df$`ICC classifier`)
df$`ICC MRCA`<- as.numeric(df$`ICC MRCA`)
df$MRCGrimwade <- as.numeric(df$MRCGrimwade)
df$t1_CR_1_Cri_2_other_3 <- as.numeric(df$t1_CR_1_Cri_2_other_3)
df$VAF_gene1_t1 <- as.numeric(df$VAF_gene1_t1)
df$VAF_gene2_t1 <- as.numeric(df$VAF_gene2_t1)
df$VAF_gene3_t1 <- as.numeric(df$VAF_gene3_t1)
df$VAF_gene4_t1 <- as.numeric(df$VAF_gene4_t1)
df$VAF_gene1_t2 <- as.numeric(df$VAF_gene1_t2)
df$VAF_gene2_t2 <- as.numeric(df$VAF_gene2_t2)
df$VAF_gene3_t2 <- as.numeric(df$VAF_gene3_t2)
df$VAF_gene4_t2 <- as.numeric(df$VAF_gene4_t2)
df$Sample_Date_MRD_t2 <- as.numeric(df$Sample_Date_MRD_t2)
df$alloTx_Dat <- as.numeric(df$alloTx_Dat)
df$VAF_X_adjusted_gene1_t1 <- as.numeric(df$VAF_X_adjusted_gene1_t1)
df$VAF_X_adjusted_gene2_t1 <- as.numeric(df$VAF_X_adjusted_gene2_t1)
df$VAF_X_adjusted_gene3_t1 <- as.numeric(df$VAF_X_adjusted_gene3_t1)
df$VAF_X_adjusted_gene4_t1 <- as.numeric(df$VAF_X_adjusted_gene4_t1)
df$VAF_X_adjusted_gene1_t2 <- as.numeric(df$VAF_X_adjusted_gene1_t2)
df$VAF_X_adjusted_gene2_t2 <- as.numeric(df$VAF_X_adjusted_gene2_t2)
df$VAF_X_adjusted_gene3_t2 <- as.numeric(df$VAF_X_adjusted_gene3_t2)
df$VAF_X_adjusted_gene4_t2 <- as.numeric(df$VAF_X_adjusted_gene4_t2)
df$t1_MRD_VAF_BCOR_X_adjusted <- as.numeric(df$t1_MRD_VAF_BCOR_X_adjusted)
df$t2_MRD_VAF_BCOR_X_adjusted <- as.numeric(df$t2_MRD_VAF_BCOR_X_adjusted)
df$t1_MRD_VAF_STAG2_X_adjusted <- as.numeric(df$t1_MRD_VAF_STAG2_X_adjusted)
df$t2_MRD_VAF_STAG2_X_adjusted <- as.numeric(df$t2_MRD_VAF_STAG2_X_adjusted)
```

Subsets for the following analysis
```{r}
t2tx=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1)
t2mpositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_MRGM_positive1_negative0==1)
t2mnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2==1 & df$t2_MRD_MRGM_positive1_negative0==0)
t1andt2=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2==1)
t1andt2mpositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2==1 & df$t1_MRD_MRGM_positive1_negative0==1)
t1andt2mnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2==1 & df$t1_MRD_MRGM_positive1_negative0==0)
MRD_MRGM=subset(df, df$`Einschluss t1_oder_t2`==1)
oldt1andt2=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1)
oldt1cohort=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1)
oldt2cohort=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1)
oldt1andt2=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1)
oldtxonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$Tx_yes_no==1)
oldnontxonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$Tx_yes_no==0)
oldt1compositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_Comutations_positive1_negative0==1)
oldt1comnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_Comutations_positive1_negative0==0)
oldt1mpositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_MRGM_positive1_negative0==1)
oldt1mnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_MRGM_positive1_negative0==0)
oldt2compositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_Comutations_positive1_negative0==1)
oldt2comnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_Comutations_positive1_negative0==0)
oldt1compositivnodta=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_Comutations_noDTA==1)
oldt1comnegativnodta=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t1_MRD_Comutations_noDTA==0)
oldt2compositivnodta=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_Comutations_noDTA==1)
oldt2comnegativnodta=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_Comutations_noDTA==0)
oldt2mpositiv=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_MRGM_positive1_negative0==1)
oldt2mnegativ=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_MRD_MRGM_positive1_negative0==0)
noNPM1=subset(df, `Einschluss t1_oder_t2`==1 & NPM1_code==0)
noTP53=subset(df, `Einschluss t1_oder_t2`==1 & TP53_code==0)
noNPM1noTP53=subset(df, `Einschluss t1_oder_t2`==1 & NPM1_code==0 & TP53_code==0)
ELNintermediate=subset(df, `Einschluss t1_oder_t2`==1 & ELN==2)
noTP53_Tx_only=subset(df, `Einschluss t1_oder_t2`==1 & TP53_code==0 & Tx_yes_no==1)
noNPM1noTP53_Tx_only=subset(df, `Einschluss t1_oder_t2`==1 & NPM1_code==0 & TP53_code==0 & Tx_yes_no==1)
overlapt1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_MRD_MRGM_positive1_negative0==1)
overlapchemoonlyt1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Type_of_consolidation_therapy==0 & df$t1_MRD_MRGM_positive1_negative0==1)
overlaptxonlyt1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & (df$Type_of_consolidation_therapy == 1 | df$Type_of_consolidation_therapy == 2) & df$t1_MRD_MRGM_positive1_negative0==1)
overlapchemotxt1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Type_of_consolidation_therapy==2 & df$t1_MRD_MRGM_positive1_negative0==1)
overlapt1posnoASXL1=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_MRD_MRGM_noASXL1==1)
t1FLT3mrdpos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t1_FLT3_MRD==1)
t1FLT3mrdneg=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t1_FLT3_MRD==0)
t2FLT3mrdpos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_FLT3_MRD==1)
t2FLT3mrdneg=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_FLT3_MRD==0)
t1NPM1mrdpos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t1_NPM1_MRD==1)
t1NPM1mrdneg=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t1_NPM1_MRD==0)
t2NPM1mrdpos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_NPM1_MRD==1)
t2NPM1mrdneg=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_NPM1_MRD==0)
oldt1cohortnonpm1=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$NPM1_code!=1)
oldt2cohortnonpm1=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$NPM1_code !=1)
overlapt1posnonpm1=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$t1_MRD_MRGM_positive1_negative0==1 & df$NPM1_code !=1)
onlyt1=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$t2_eingeschlossen1_ausgeschlossen0!=1)
onlyt2=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`!=1 & df$t2_eingeschlossen1_ausgeschlossen0==1)
t1txonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$Tx_yes_no==1)
t2txonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1)
t1chemoonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$Type_of_consolidation_therapy==0)
t2chemoonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Type_of_consolidation_therapy==0)
t1pbonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$Sample_Material_PB_1_BM_2==1)
t1bmonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$`t1_eingeschlossen 1 _ ausgeschlossen 0`==1 & df$Sample_Material_PB_1_BM_2==2)
t2pbonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Sample_Material_t2_PB_1_BM_2==1)
t2bmonly=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Sample_Material_t2_PB_1_BM_2==2)
```

Table 1: Clinical and transplantation-associated characteristics of patients, compared between MRD-positive and MRD-negative patients based on MRG mutations
Patient characteristics of all transplanted patients included in t2 
(main cohort / t2tx)
```{r}
nrow(t2tx)
#age
summary(t2tx$Age_at_ED)
summary(t2mpositiv$Age_at_ED)
summary(t2mnegativ$Age_at_ED)
wilcox.test(t2mpositiv$Age_at_ED, t2mnegativ$Age_at_ED)
#sex
table(t2tx$Sex, t2tx$t2_MRD_MRGM_positive1_negative0)
table_sex_t2<-table(t2tx$Sex, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_sex_t2)
#ICC classifier
table(t2tx$`ICC classifier`, t2tx$t2_MRD_MRGM_positive1_negative0)
table_icc_classifier<-table(t2tx$`ICC classifier`, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$`ICC classifier`)])
chisq.test(table_icc_classifier)
fisher.test(table_icc_classifier) #Fisher-Test weil Werte <5? 
#AML_type
table(t2tx$AML_Type, t2tx$t2_MRD_MRGM_positive1_negative0)
table_AMLtype_t2<-table(t2tx$AML_Type, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_AMLtype_t2)
fisher.test(table_AMLtype_t2) #Fisher-Test weil Werte <5? 
#ecog
table(t2tx$ECOG_Code, t2tx$t2_MRD_MRGM_positive1_negative0)
table_ECOG_t2<-table(t2tx$ECOG_Code, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_ECOG_t2)
fisher.test(table_ECOG_t2) #Fisher-Test weil Werte <5? 
#karyotype
table(t2tx$KT_complex, t2tx$t2_MRD_MRGM_positive1_negative0)
table_KTcomplex_t2<-table(t2tx$KT_complex, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_KTcomplex_t2)
fisher.test(table_KTcomplex_t2) #Fisher-Test weil Werte <5? 
#MRC Grimwade
table(t2tx$MRCGrimwade, t2tx$t2_MRD_MRGM_positive1_negative0)
table_grimwade_t2<-table(t2tx$MRCGrimwade, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$MRCGrimwade)])
chisq.test(table_grimwade_t2)
fisher.test(table_grimwade_t2) #Fisher-Test weil Werte <5? 
#MRCA ICC
table(t2tx$`ICC MRCA`, t2tx$t2_MRD_MRGM_positive1_negative0)
table_MRCA_t2<-table(t2tx$`ICC MRCA`, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$`ICC MRCA`)])
chisq.test(table_MRCA_t2)
fisher.test(table_MRCA_t2) #Fisher-Test weil Werte <5? 
#wbc
summary(t2tx$WBC_at_ED)
summary(t2mpositiv$WBC_at_ED)
summary(t2mnegativ$WBC_at_ED)
wilcox.test(t2mpositiv$WBC_at_ED, t2mnegativ$WBC_at_ED)
#Hgb
summary(t2tx$Hb_at_ED)
summary(t2mpositiv$Hb_at_ED)
summary(t2mnegativ$Hb_at_ED)
wilcox.test(t2mpositiv$Hb_at_ED, t2mnegativ$Hb_at_ED)
#Plt
summary(t2tx$Platelets_at_ED)
summary(t2mpositiv$Platelets_at_ED)
summary(t2mnegativ$Platelets_at_ED)
wilcox.test(t2mpositiv$Platelets_at_ED, t2mnegativ$Platelets_at_ED)
```
Table of transplantation-associated characteristics of all pts included in t2
(main cohort / t2tx)
```{r}
nrow(t2tx)
#HCT-CI score before Tx
t2tx$HCTCI_over_under_2 <- ifelse(t2tx$HCTCI_at_ED <= 2, 0, 1)
table(t2tx$HCTCI_over_under_2, t2tx$t2_MRD_MRGM_positive1_negative0)
table_hctci<-table(t2tx$HCTCI_over_under_2, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$HCTCI_over_under_2)])
chisq.test(table_hctci)
fisher.test(table_hctci) #Fisher-Test weil Werte <5? 
#number of chemotherapy cycles before Tx
table(t2tx$cycles_before_Tx_1_2_3ormore, t2tx$t2_MRD_MRGM_positive1_negative0)
table_cyclesbeforeTx<-table(t2tx$cycles_before_Tx_1_2_3ormore, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_cyclesbeforeTx)
#Donor match
t2tx$donor_match_grouped <- ifelse(t2tx$Donor_match_Tx1 == 4, 3, t2tx$Donor_match_Tx1)
table(t2tx$donor_match_grouped, t2tx$t2_MRD_MRGM_positive1_negative0)
table_donor_match<-table(t2tx$donor_match_grouped, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$donor_match_grouped)])
chisq.test(table_donor_match)
#Conditioning therapy MAC/RIC
table(t2tx$MAC_vs_RIC, t2tx$t2_MRD_MRGM_positive1_negative0)
table_conditioning<-table(t2tx$MAC_vs_RIC, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_conditioning)
fisher.test(table_conditioning) #Fisher-Test weil Werte <5? 
#Stem cell source
table(t2tx$StemCell_Source, t2tx$t2_MRD_MRGM_positive1_negative0)
table_stemcellsource<-table(t2tx$StemCell_Source, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$StemCell_Source)])
chisq.test(table_stemcellsource)
fisher.test(table_stemcellsource) #Fisher-Test weil Werte <5? 
#Donor_sex
table(t2tx$Donor_Sex_atTx1, t2tx$t2_MRD_MRGM_positive1_negative0)
table_donorsex<-table(t2tx$Donor_Sex_atTx1, t2tx$t2_MRD_MRGM_positive1_negative0)
chisq.test(table_donorsex)
#CMV status
table(t2tx$CMVstatus_PnegDneg_vs_rest, t2tx$t2_MRD_MRGM_positive1_negative0)
table_cmvstatus<-table(t2tx$CMVstatus_PnegDneg_vs_rest, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$CMVstatus_PnegDneg_vs_rest)])
chisq.test(table_cmvstatus)
#CMV reactivation
# missing!

#aGvHD
t2tx$aGvHD_grouped <- ifelse(t2tx$aGvHDmax_grade == 0, 0, ifelse(t2tx$aGvHDmax_grade %in% c(1, 2), 1, 2))
table(t2tx$aGvHD_grouped, t2tx$t2_MRD_MRGM_positive1_negative0)
table_agvhd_grade<-table(t2tx$aGvHD_grouped, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$aGvHD_grouped)])
chisq.test(table_agvhd_grade)
#cGvHD
table(t2tx$cGvHD_Tx1_grade, t2tx$t2_MRD_MRGM_positive1_negative0)
table_cgvhd<-table(t2tx$cGvHD_Tx1_grade, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$cGvHD_Tx1_grade)])
chisq.test(table_cgvhd)
```


Table 2: MRD characteristics in patients with and without detectable MRD of MRG mutations
MRD characteristics t2 (Median VAF, range)
```{r}
#ASXL1
table (t2tx$t2_MRD_ASXL1)
table(t2tx$t2_MRD_ASXL1, t2tx$t2_MRD_MRGM_positive1_negative0)
table_asxl1<-table(t2tx$t2_MRD_ASXL1, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_ASXL1)])
summary(t2tx$t2_MRD_VAF_ASXL1)
chisq.test(table_asxl1)
fisher.test(table_asxl1) #Fisher-Test weil Werte <5? 
#BCOR
table (t2tx$t2_MRD_BCOR)
table(t2tx$t2_MRD_BCOR, t2tx$t2_MRD_MRGM_positive1_negative0)
table_bcor<-table(t2tx$t2_MRD_BCOR, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_BCOR)])
summary(t2tx$t2_MRD_VAF_BCOR_X_adjusted)
chisq.test(table_bcor)
fisher.test(table_bcor) #Fisher-Test weil Werte <5? 
#EZH2
table(t2tx$t2_MRD_EZH2)
table(t2tx$t2_MRD_EZH2, t2tx$t2_MRD_MRGM_positive1_negative0)
table_ezh2<-table(t2tx$t2_MRD_EZH2, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_EZH2)])
summary(t2tx$t2_MRD_VAF_EZH2)
chisq.test(table_ezh2)
fisher.test(table_ezh2) #Fisher-Test weil Werte <5? 
#RUNX1
table(t2tx$t2_MRD_RUNX1)
table(t2tx$t2_MRD_RUNX1, t2tx$t2_MRD_MRGM_positive1_negative0)
table_runx1<-table(t2tx$t2_MRD_RUNX1, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_RUNX1)])
summary(t2tx$t2_MRD_VAF_RUNX1)
chisq.test(table_runx1)
fisher.test(table_runx1) #Fisher-Test weil Werte <5? 
#SF3B1
table(t2tx$t2_MRD_SF3B1)
table(t2tx$t2_MRD_SF3B1, t2tx$t2_MRD_MRGM_positive1_negative0)
table_sf3b1<-table(t2tx$t2_MRD_SF3B1, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_SF3B1)])
summary(t2tx$t2_MRD_VAF_SF3B1)
chisq.test(table_sf3b1)
#fisher.test(table_sf3b1) #Fisher-Test weil Werte <5? 
#SRSF2
table(t2tx$t2_MRD_SRSF2)
table(t2tx$t2_MRD_SRSF2, t2tx$t2_MRD_MRGM_positive1_negative0)
table_srsf2<-table(t2tx$t2_MRD_SRSF2, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_SRSF2)])
summary(t2tx$t2_MRD_VAF_SRSF2)
chisq.test(table_srsf2)
fisher.test(table_srsf2) #Fisher-Test weil Werte <5? 
#STAG2
table(t2tx$t2_MRD_STAG2)
table(t2tx$t2_MRD_STAG2, t2tx$t2_MRD_MRGM_positive1_negative0)
table_stag2<-table(t2tx$t2_MRD_STAG2, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_STAG2)])
summary(t2tx$t2_MRD_VAF_STAG2_X_adjusted)
chisq.test(table_stag2)
#fisher.test(table_stag2) #Fisher-Test weil Werte <5? 
#U2AF1
table(t2tx$t2_MRD_U2AF1)
table(t2tx$t2_MRD_U2AF1, t2tx$t2_MRD_MRGM_positive1_negative0)
table_u2af1<-table(t2tx$t2_MRD_U2AF1, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_U2AF1)])
summary(t2tx$t2_MRD_VAF_U2AF1)
chisq.test(table_u2af1)
fisher.test(table_u2af1) #Fisher-Test weil Werte <5? 
#ZRSR2
table(t2tx$t2_MRD_ZRSR2)
table(t2tx$t2_MRD_ZRSR2, t2tx$t2_MRD_MRGM_positive1_negative0)
table_zrsr2<-table(t2tx$t2_MRD_ZRSR2, t2tx$t2_MRD_MRGM_positive1_negative0)
table(t2tx$t2_MRD_MRGM_positive1_negative0[is.na(t2tx$t2_MRD_ZRSR2)])
summary(t2tx$t2_MRD_VAF_ZRSR2)
#chisq.test(table_zrsr2)
#fisher.test(table_zrsr2) #Fisher-Test weil Werte <5? 

#VAF Median and Range, only of MRD positives:
#subsets
t2asxl1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_ASXL1 ==1)
t2bcorpos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_BCOR ==1)
t2EZH2pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_EZH2 ==1)
t2RUNX1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_RUNX1 ==1)
t2SF3B1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_SF3B1 ==1)
t2SRSF2pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_SRSF2 ==1)
t2STAG2pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_STAG2 ==1)
t2U2AF1pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_U2AF1 ==1)
t2ZRSR2pos=subset(df, df$`Einschluss t1_oder_t2`==1 & df$t2_eingeschlossen1_ausgeschlossen0==1 & df$Tx_yes_no==1 & df$Txin_CR1_vs_CR2 == 1 & df$t2_MRD_ZRSR2 ==1)
#VAF Median & Range
summary(t2asxl1pos$t2_MRD_VAF_ASXL1)
summary(t2bcorpos$t2_MRD_VAF_BCOR_X_adjusted)
summary(t2EZH2pos$t2_MRD_VAF_EZH2)
summary(t2RUNX1pos$t2_MRD_VAF_RUNX1)
summary(t2SF3B1pos$t2_MRD_VAF_SF3B1)
summary(t2SRSF2pos$t2_MRD_VAF_SRSF2)
summary(t2STAG2pos$t2_MRD_VAF_STAG2_X_adjusted)
summary(t2U2AF1pos$t2_MRD_VAF_U2AF1)
summary(t2ZRSR2pos$t2_MRD_VAF_ZRSR2)
```

Figure 1: Frequency of MRG mutations and MRD persistence 30-180 days after alloHCT
Mutational landscape: Barplot -> Graphpad
P-Values for comparing probability of positivity in t1 and t2
```{r}
#ASXL1
table_t1t2_ASXL1 <- table(t1andt2$t1_MRD_ASXL1, t1andt2$t2_MRD_ASXL1)
table_t1t2_ASXL1
mcnemar.test(table_t1t2_ASXL1)

all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_ASXL1, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_ASXL1, levels = all_levels)
table_t1t2_ASXL1_square <- table(t1_factor, t2_factor)
print(table_t1t2_ASXL1_square)
mcnemar.test(table_t1t2_ASXL1_square)

#BCOR
#table_t1t2_BCOR <- table(t1andt2$t1_MRD_BCOR, t1andt2$t2_MRD_BCOR)
#table_t1t2_BCOR
#mcnemar.test(table_t1t2_BCOR)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_BCOR, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_BCOR, levels = all_levels)
table_t1t2_BCOR_square <- table(t1_factor, t2_factor)
print(table_t1t2_BCOR_square)
mcnemar.test(table_t1t2_BCOR_square)

#EZH2
#table_t1t2_EZH2 <- table(t1andt2$t1_MRD_EZH2, t1andt2$t2_MRD_EZH2)
#table_t1t2_EZH2
#mcnemar.test(table_t1t2_EZH2)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_EZH2, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_EZH2, levels = all_levels)
table_t1t2_EZH2_square <- table(t1_factor, t2_factor)
print(table_t1t2_EZH2_square)
mcnemar.test(table_t1t2_EZH2_square)

#RUNX1
table_t1t2_RUNX1 <- table(t1andt2$t1_MRD_RUNX1, t1andt2$t2_MRD_RUNX1)
table_t1t2_RUNX1
mcnemar.test(table_t1t2_RUNX1)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_RUNX1, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_RUNX1, levels = all_levels)
table_t1t2_RUNX1_square <- table(t1_factor, t2_factor)
print(table_t1t2_RUNX1_square)
mcnemar.test(table_t1t2_RUNX1_square)

#SF3B1
#table_t1t2_SF3B1 <- table(t1andt2$t1_MRD_SF3B1, t1andt2$t2_MRD_SF3B1)
#table_t1t2_SF3B1
#mcnemar.test(table_t1t2_SF3B1)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_SF3B1, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_SF3B1, levels = all_levels)
table_t1t2_SF3B1_square <- table(t1_factor, t2_factor)
print(table_t1t2_SF3B1_square)
mcnemar.test(table_t1t2_SF3B1_square)

#SRSF2
table_t1t2_SRSF2 <- table(t1andt2$t1_MRD_SRSF2, t1andt2$t2_MRD_SRSF2)
table_t1t2_SRSF2
mcnemar.test(table_t1t2_SRSF2)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_SRSF2, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_SRSF2, levels = all_levels)
table_t1t2_SRSF2_square <- table(t1_factor, t2_factor)
print(table_t1t2_SRSF2_square)
mcnemar.test(table_t1t2_SRSF2_square)


#STAG2
#table_t1t2_STAG2 <- table(t1andt2$t1_MRD_STAG2, t1andt2$t2_MRD_STAG2)
#table_t1t2_STAG2
#mcnemar.test(table_t1t2_STAG2)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_STAG2, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_STAG2, levels = all_levels)
table_t1t2_STAG2_square <- table(t1_factor, t2_factor)
print(table_t1t2_STAG2_square)
mcnemar.test(table_t1t2_STAG2_square)

#U2AF1
#table_t1t2_U2AF1 <- table(t1andt2$t1_MRD_U2AF1, t1andt2$t2_MRD_U2AF1)
#table_t1t2_U2AF1
#mcnemar.test(table_t1t2_U2AF1)
all_levels <- c(0, 1) 
t1_factor <- factor(t1andt2$t1_MRD_U2AF1, levels = all_levels)
t2_factor <- factor(t1andt2$t2_MRD_U2AF1, levels = all_levels)
table_t1t2_U2AF1_square <- table(t1_factor, t2_factor)
print(table_t1t2_U2AF1_square)
mcnemar.test(table_t1t2_U2AF1_square)

#ZRSR2
table_t1t2_ZRSR2 <- table(t1andt2$t1_MRD_ZRSR2, t1andt2$t2_MRD_ZRSR2)
table_t1t2_ZRSR2
mcnemar.test(table_t1t2_ZRSR2)
```

Median Follow-Up (OS_fromTx_years or OS_fromTx_years_censored?)
```{r}

fit_rev <- survfit(Surv(OS_fromTx_years_censored, 1 - OS_status) ~ 1, data = t2tx)
summary(fit_rev)$table["median"]

#range
follow_up_range <- range(t2tx$OS_fromTx_years_censored)
cat("Min", follow_up_range[1], "\n")
cat("Max", follow_up_range[2], "\n")
```

Figure 2: Overall survival, relapse-free survival and cumulative incidence of relapse for MRG mutation MRD-positive and MRD-negative patients after alloHCT 
OS t2tx (calculated from Tx) (updated figure design)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure2_OS_t2tx_v6.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
RFS t2tx (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure2_RFS_t2tx_v6.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_MRGM_positive1_negative0, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
CIR t2tx (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure2_CIR_t2tx_v6.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t2tx$RFS_fromTx_years_censored, fstatus = t2tx$CIR_event, group = t2tx$t2_MRD_MRGM_positive1_negative0)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
fig$data$combined_label <- factor(fig$data$combined_label,
                                  levels = c("1_1", "0_1", "1_2", "0_2"))

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )
#risk table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_MRGM_positive1_negative0, data = t2tx)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t2tx,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),

    legend.position = c(0.5, 1.005), #bottom center
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame of legend
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6) # using inches
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

#P-Wert CIR
fine_gray_test_cir <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 1, t2tx$t2_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_cir)

#P-Wert NRM
fine_gray_test_nrm <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 2, t2tx$t2_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_nrm)
```
`

Table 4: Multivate Analysis of all patients after alloHCT
Define new variables for the multivariate analysis
Age by decades
```{r}
t2tx$age_cat <- case_when(t2tx$Age_at_ED <= 20 ~ '1',
                  between(t2tx$Age_at_ED, 20.1 , 30) ~ '2',
                  between(t2tx$Age_at_ED, 30.1, 40) ~ '3',
                  between(t2tx$Age_at_ED, 40.1, 50) ~ '4',
                  between(t2tx$Age_at_ED, 50.1, 60) ~ '5',
                  between(t2tx$Age_at_ED, 60.1, 70) ~ '6',
                  t2tx$Age_at_ED >=70.1 ~ '7')
head(t2tx$age_cat)

```
Log transform Hb, Plt and WBC
```{r}
t2tx$Hb_log<-log2(t2tx$Hb_at_ED)
t2tx$WBC_log<-log2(t2tx$WBC_at_ED)
t2tx$Plt_log<-log2(t2tx$Platelets_at_ED)
```
MVA of OS step down
```{r}
library(mice)
t2cohort_imp<-t2tx%>%select(OS_status, OS_fromTx_years_censored, Hb_at_ED, WBC_at_ED, Platelets_at_ED, Age_at_ED,KT_complex, NPM1_code, t2_MRD_MRGM_positive1_negative0, Sex, AML_Type)

init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_at_ED")]="pmm" 
meth[c("WBC_at_ED")]="pmm" 
meth[c("Platelets_at_ED")]="pmm"
meth[c("Age_at_ED")]="pmm"
meth[c("KT_complex")]="logreg"
meth[c("OS_status")]=""
meth[c("OS_fromTx_years_censored")]=""
meth[c("t2_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_at_ED + WBC_at_ED +Platelets_at_ED + Age_at_ED  + KT_complex +imputed$NPM1_code + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Platelets_at_ED + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
MVA step up
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_at_ED , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~WBC_at_ED , data = imputed) %>% tbl_regression (exp = TRUE)

```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Platelets_at_ED, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~Age_at_ED , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ KT_complex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~NPM1_code , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
This results in the same model with platelets and MRG mutations.
MVA with age in decades and log transformed Hb, WBC and platelets.
```{r}
t2tx$age_cat<-as.numeric(t2tx$age_cat)
t2cohort_imp<-t2tx%>%select(OS_status, OS_fromTx_years_censored, Hb_log, WBC_log, Plt_log, age_cat,KT_complex, NPM1_code, t2_MRD_MRGM_positive1_negative0, Sex, AML_type_CODE)

init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_log")]="pmm" 
meth[c("WBC_log")]="pmm" 
meth[c("Plt_log")]="pmm"
meth[c("age_cat")]="logreg"
meth[c("KT_complex")]="logreg"
meth[c("OS_status")]=""
meth[c("OS_fromTx_years_censored")]=""
meth[c("t2_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"
meth[c("Sex")]="logreg"
meth[c("AML_type_CODE")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_log + WBC_log +Plt_log + age_cat  + KT_complex +imputed$NPM1_code + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Plt_log  + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ WBC_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~Plt_log, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~age_cat, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ KT_complex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~imputed$NPM1_code , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ Sex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ AML_type_CODE, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~ Plt_log + t2_MRD_MRGM_positive1_negative0 + Sex, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~ Plt_log + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```

MVA RFS
```{r}
t2cohort_imp<-t2tx%>%select(RFS_status, RFS_fromTx_years_censored, Hb_at_ED, WBC_at_ED, Platelets_at_ED, Age_at_ED,KT_complex, NPM1_code, t2_MRD_MRGM_positive1_negative0, Sex, AML_type_CODE)
t2cohort_imp<-t2cohort_imp%>%drop_na(RFS_fromTx_years_censored)
init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_at_ED")]="pmm" 
meth[c("WBC_at_ED")]="pmm" 
meth[c("Platelets_at_ED")]="pmm"
meth[c("Age_at_ED")]="pmm"
meth[c("KT_complex")]="logreg"
meth[c("RFS_status")]=""
meth[c("RFS_fromTx_years_censored")]=""
meth[c("t2_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"
meth[c("Sex")]="logreg"
meth[c("AML_type_CODE")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Hb_at_ED + WBC_at_ED +Platelets_at_ED + Age_at_ED  + KT_complex +imputed$NPM1_code + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
RFS with age in decades and log transformed
```{r}
t2tx$age_cat<-as.numeric(t2tx$age_cat)
t2cohort_imp<-t2tx%>%select(RFS_status, RFS_fromTx_years_censored, Hb_log, WBC_log, Plt_log, age_cat,KT_complex, NPM1_code, t2_MRD_MRGM_positive1_negative0, Sex, AML_type_CODE)
t2cohort_imp<-t2cohort_imp%>%drop_na(RFS_fromTx_years_censored)

init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_log")]="pmm" 
meth[c("WBC_log")]="pmm" 
meth[c("Plt_log")]="pmm"
meth[c("age_cat")]="logreg"
meth[c("KT_complex")]="logreg"
meth[c("RFS_status")]=""
meth[c("RFS_fromTx_years_censored")]=""
meth[c("t2_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"
meth[c("AML_type_CODE")]="logreg"
meth[c("Sex")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Hb_log + WBC_log +Plt_log + age_cat  + KT_complex +imputed$NPM1_code + t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
With the log transformed approach the MVA is not significant in RFS.
MVA step down
```{r}
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Hb_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ WBC_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Plt_log, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~age_cat, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ KT_complex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~imputed$NPM1_code , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ t2_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ Sex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ AML_type_CODE, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Plt_log+Sex, data = imputed) %>% tbl_regression (exp = TRUE)
```


Figure 3: MRD of MRG mutations without ASXL1 shows no prognostic value in OS, RFS and CIR

OS t2tx without ASXL1 (calculated from Tx) (updated figure design)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure3_OS_withoutASXL1_t2tx_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_MRGM_noASXL1, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_MRGM_noASXL1, data=t2tx)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t2tx$t2_MRD_MRGM_noASXL1, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
RFS t2tx without ASXL1 (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure3_RFS_withoutASXL1_t2tx_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_MRGM_noASXL1, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_MRGM_noASXL1, data=t2tx)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_MRGM_noASXL1, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
CIR t2tx without ASXL1 (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure3_CIR_withoutASXL1_t2tx_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t2tx$RFS_fromTx_years_censored, fstatus = t2tx$CIR_event, group = t2tx$t2_MRD_MRGM_noASXL1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
fig$data$combined_label <- factor(fig$data$combined_label,
                                  levels = c("1_1", "0_1", "1_2", "0_2"))

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )
#risk table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_MRGM_noASXL1, data = t2tx)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t2tx,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),

    legend.position = c(0.5, 1.005), #bottom center
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame of legend
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6) # using inches
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

#P-Wert CIR
fine_gray_test_cir <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 1, t2tx$t2_MRD_MRGM_noASXL1)
summary(fine_gray_test_cir)

#P-Wert NRM
fine_gray_test_nrm <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 2, t2tx$t2_MRD_MRGM_noASXL1)
summary(fine_gray_test_nrm)
```


Table 5: Multivate Analysis of patients assessed for MRD of MRG mutations after induction chemotherapy
MVA
```{r}
t1andt2$age_cat <- case_when(t1andt2$Age_at_ED <= 20 ~ '1',
                  between(t1andt2$Age_at_ED, 20.1 , 30) ~ '2',
                  between(t1andt2$Age_at_ED, 30.1, 40) ~ '3',
                  between(t1andt2$Age_at_ED, 40.1, 50) ~ '4',
                  between(t1andt2$Age_at_ED, 50.1, 60) ~ '5',
                  between(t1andt2$Age_at_ED, 60.1, 70) ~ '6',
                  t1andt2$Age_at_ED >=70.1 ~ '7')
head(t1andt2$age_cat)

```
Log transform Hb, Plt and WBC
```{r}
t1andt2$Hb_log<-log2(t1andt2$Hb_at_ED)
t1andt2$WBC_log<-log2(t1andt2$WBC_at_ED)
t1andt2$Plt_log<-log2(t1andt2$Platelets_at_ED)
```
```{r}
t1andt2$age_cat<-as.numeric(t1andt2$age_cat)
t2cohort_imp<-t1andt2%>%select(OS_status, OS_fromTx_years_censored, Hb_log, WBC_log, Plt_log, age_cat,KT_complex, NPM1_code, t1_MRD_MRGM_positive1_negative0, Sex, AML_type_CODE)

init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_log")]="pmm" 
meth[c("WBC_log")]="pmm" 
meth[c("Plt_log")]="pmm"
meth[c("age_cat")]="logreg"
meth[c("KT_complex")]="logreg"
meth[c("OS_status")]=""
meth[c("OS_fromTx_years_censored")]=""
meth[c("t1_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"
meth[c("Sex")]="logreg"
meth[c("AML_type_CODE")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_log + WBC_log +Plt_log + age_cat  + KT_complex +imputed$NPM1_code + t1_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```

```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~Hb_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ WBC_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~Plt_log, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~age_cat, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ KT_complex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~imputed$NPM1_code , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ t1_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ Sex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(OS_fromTx_years_censored, OS_status)~ AML_type_CODE, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}
coxph(Surv(OS_fromTx_years_censored, OS_status)~ Plt_log+Sex+AML_type_CODE, data = imputed) %>% tbl_regression (exp = TRUE)
```
```{r}

t2cohort_imp<-t1andt2%>%select(RFS_status, RFS_fromTx_years_censored, Hb_log, WBC_log, Plt_log, age_cat,KT_complex, NPM1_code,t1_MRD_MRGM_positive1_negative0 , Sex, AML_type_CODE)
t2cohort_imp<-t2cohort_imp%>%drop_na(RFS_fromTx_years_censored)

init = mice(t2cohort_imp, maxit=0) 
meth = init$method
predM = init$predictorMatrix

meth[c("Hb_log")]="pmm" 
meth[c("WBC_log")]="pmm" 
meth[c("Plt_log")]="pmm"
meth[c("age_cat")]="logreg"
meth[c("KT_complex")]="logreg"
meth[c("RFS_status")]=""
meth[c("RFS_fromTx_years_censored")]=""
meth[c("t1_MRD_MRGM_positive1_negative0")]=""
meth[c("NPM1_code")]="logreg"
meth[c("AML_type_CODE")]="logreg"
meth[c("Sex")]="logreg"

set.seed(123)
t2cohort_imp<-(t2cohort_imp%>% mice( m=5, method=meth, predictorMatrix=predM, print = FALSE))
imputed<-complete(t2cohort_imp)
sapply(imputed, function(x) sum(is.na(x)))

coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Hb_log + WBC_log +Plt_log + age_cat  + KT_complex +imputed$NPM1_code + t1_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
```
MVA step down
```{r}
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Hb_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ WBC_log , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~Plt_log, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~age_cat, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ KT_complex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~imputed$NPM1_code , data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ t1_MRD_MRGM_positive1_negative0, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ Sex, data = imputed) %>% tbl_regression (exp = TRUE)
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ AML_type_CODE, data = imputed) %>% tbl_regression (exp = TRUE)
```

```{r}
coxph(Surv(RFS_fromTx_years_censored, RFS_status)~ AML_type_CODE+Sex+Plt_log, data = imputed) %>% tbl_regression (exp = TRUE)
```



Figure 4: MRD status of MRG mutations does not predict overall survival, relapse-free survival or cumulative incidence of relapse after induction chemotherapy.
OS t1andt2
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure4_OS_t1andt2_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
RFS t1andt2
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure4_RFS_t1andt2_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_positive1_negative0, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
(CIR t1andt2 original)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure4_CIR_t1andt2_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_MRGM_positive1_negative0)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
fig$data$combined_label <- factor(fig$data$combined_label,
                                  levels = c("1_1", "0_1", "1_2", "0_2"))

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )
#risk table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_positive1_negative0, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),

    legend.position = c(0.5, 1.005), #bottom center
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame of legend
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6) # using inches
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

#P-Wert CIR
fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_cir)

#P-Wert NRM
fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_nrm)
```
CIR t1andt2 (changed order overlapping lines)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_Figure4_CIR_t1andt2_v3.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_MRGM_positive1_negative0)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )

# Risk Table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_positive1_negative0, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_MRGM_positive1_negative0)
summary(fine_gray_test_nrm)
```


+++ SUPPLEMENT +++

Supplementary Figure 1:
CONSORT diagram

Supplementary Figure 2:
OS t1andt2 without ASXL1
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS2_OS_t1andt2_noASXL1_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
RFS t1andt2 without ASXL1
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS2_RFS_t1andt2_noASXL1_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_noASXL1, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
(CIR t1andt2  without ASXL1 original)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS2_CIR_t1andt2_noASXL1_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_MRGM_noASXL1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
fig$data$combined_label <- factor(fig$data$combined_label,
                                  levels = c("1_1", "0_1", "1_2", "0_2"))

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )
#risk table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_noASXL1, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),

    legend.position = c(0.5, 1.005), #bottom center
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame of legend
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6) # using inches
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

#P-Wert CIR
fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_MRGM_noASXL1)
summary(fine_gray_test_cir)

#P-Wert NRM
fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_MRGM_noASXL1)
summary(fine_gray_test_nrm)
```
CIR t1andt2 without ASXL1 (changed order overlapping lines)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS2_CIR_t1andt2_noASXL1_v3.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_MRGM_noASXL1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )

# Risk Table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_MRGM_noASXL1, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_MRGM_noASXL1)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_MRGM_noASXL1)
summary(fine_gray_test_nrm)
```

Supplementary Figure 3: 
t1andt2 RUNX1 only
OS t1andt2 RUNX1
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS3_OS_t1andt2_RUNX1only_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_RUNX1, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$t1_MRD_RUNX1, data=t1andt2)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t1andt2$t1_MRD_RUNX1, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
RFS t1andt2 RUNX1 only
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS3_RFS_t1andt2_RUNX1only_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_RUNX1, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    legend.background = element_blank() #no white frame
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$t1_MRD_RUNX1, data=t1andt2)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_RUNX1, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
(CIR t1andt2 RUNX1 only (lines overlap))
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS3_CIR_t1andt2_RUNX1only_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_RUNX1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )

# Risk Table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_RUNX1, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_RUNX1)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_RUNX1)
summary(fine_gray_test_nrm)
```
CIR t1andt2 RUNX1 only (dotted yellow line)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS3_CIR_t1andt2_RUNX1only_v3.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$t1_MRD_RUNX1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label, linetype = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2, override.aes = list(linetype = "solid"))
  ) +
  scale_linetype_manual(
    values = c(
      "0_1" = "dashed",
      "0_2" = "solid",
      "1_1" = "solid",
      "1_2" = "solid"
    ),
    guide = "none"
  )

fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$t1_MRD_RUNX1, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$t1_MRD_RUNX1)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$t1_MRD_RUNX1)
summary(fine_gray_test_nrm)
```


Supplementary Figure 4:
Line graph of VAF dynamics

Supplementary Figure 5: Conversion to MRD negativity does not predict clinical outcome
definition of converters:
```{r}
df$MRD_MRGM_conversion <- with(df, ifelse(is.na(t1_MRD_MRGM_positive1_negative0) | is.na(t2_MRD_MRGM_positive1_negative0), NA,   # Falls NA in t1 oder t2  NA
  ifelse(t1_MRD_MRGM_positive1_negative0 == 1 & t2_MRD_MRGM_positive1_negative0 == 0, 1, # Konvertiert (1  0)
  ifelse(t1_MRD_MRGM_positive1_negative0 == 1 & t2_MRD_MRGM_positive1_negative0 == 1, 0, # Nicht konvertiert (1  1)
  ifelse(t1_MRD_MRGM_positive1_negative0 == 0 & t2_MRD_MRGM_positive1_negative0 == 0, NA, # Immer negativ (0  0)  NA
  0))))) # brige Flle (z. B. 0  1) = 0

df$MRD_Comutations_conversion <- with(df, ifelse(
  is.na(t1_MRD_Comutations_noDTA) | is.na(t2_MRD_Comutations_noDTA), NA,   # Falls NA in t1 oder t2  NA
  ifelse(t1_MRD_Comutations_noDTA == 1 & t2_MRD_Comutations_noDTA == 0, 1, # Konvertiert (1  0)
  ifelse(t1_MRD_Comutations_noDTA == 1 & t2_MRD_Comutations_noDTA == 1, 0, # Nicht konvertiert (1  1)
  ifelse(t1_MRD_Comutations_noDTA == 0 & t2_MRD_Comutations_noDTA == 0, NA, # Immer negativ (0  0)  NA
  0))))) # brige Flle (z. B. 0  1) = 0

```
new definition of subsets necessary!

survival analyses of MRGM converters vs non-converters
OS MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS5_OS_t1andt2_MRGM_Conversion_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$MRD_MRGM_conversion, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-conv-", "MRD-conv+"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$MRD_MRGM_conversion, data=t1andt2)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t1andt2$MRD_MRGM_conversion, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
RFS MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS5_RFS_t1andt2_MRGM_Conversion_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$MRD_MRGM_conversion, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-conv-", "MRD-conv+"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$MRD_MRGM_conversion, data=t1andt2)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$MRD_MRGM_conversion, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
CIR MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS5_CIR_t1andt2_MRGM_Conversion_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$MRD_MRGM_conversion)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-conv-",
      "0_2" = "NRM MRD-conv-",
      "1_1" = "CIR MRD-conv+",
      "1_2" = "NRM MRD-conv+"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )

# Risk Table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$MRD_MRGM_conversion, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("no MRD conversion", "MRD conversion")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$MRD_MRGM_conversion)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$MRD_MRGM_conversion)
summary(fine_gray_test_nrm)
```


survival analyses of Comutation converters vs non-converters
OS MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS6_OS_t1andt2_comutations_Conversion_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$MRD_Comutations_conversion, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("no MRD-conv", "MRD-conv "),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t1andt2$MRD_Comutations_conversion, data=t1andt2)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t1andt2$MRD_Comutations_conversion, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
RFS MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS6_RFS_t1andt2_Comutations_Conversion_v2.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t1andt2<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$MRD_Comutations_conversion, data=t1andt2)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t1andt2,
  data = t1andt2,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("no MRD-conv", "MRD-conv "),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t1andt2$MRD_Comutations_conversion, data=t1andt2)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$MRD_Comutations_conversion, data=t1andt2) %>% 
  tbl_regression(exp = TRUE)
```
CIR MRGM converted vs not-converted patients (MRGM)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_FigureS6_CIR_t1andt2_Comutations_Conversion_v2.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t1andt2$RFS_fromTx_years_censored, fstatus = t1andt2$CIR_event, group = t1andt2$MRD_Comutations_conversion)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")
draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5, linetype = "solid") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"), 
    labels = c(
      "0_1" = "CIR MRD-conv-",
      "0_2" = "NRM MRD-conv-",
      "1_1" = "CIR MRD-conv+",
      "1_2" = "NRM MRD-conv+"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2)
  )

# Risk Table
fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t1andt2$MRD_Comutations_conversion, data = t1andt2)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t1andt2,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("no MRD conversion", "MRD conversion")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 1, t1andt2$MRD_Comutations_conversion)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t1andt2$RFS_fromTx_years_censored, t1andt2$CIR_event == 2, t1andt2$MRD_Comutations_conversion)
summary(fine_gray_test_nrm)
```

additional analysis: ASXL1 only
OS t2tx (calculated from Tx) (updated figure design)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_t2tx_ASXL1only_OS_t2tx_v1.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_ASXL1, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE, #p-value of log-rank test
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE, # confidence intervalls
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), 
  xlab = "Time from alloHCT (years)",
  ylab = "Overall survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(OS_fromTx_years_censored, OS_status)~t2tx$t2_MRD_ASXL1, data=t2tx)
coxph(Surv(OS_fromTx_years_censored, OS_status) ~ t2tx$t2_MRD_ASXL1, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
RFS t2tx (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_t2tx_ASXL1only_RFS_t2tx_v1.pdf"

dir.create(dirname(output_path), showWarnings = FALSE, recursive = TRUE)

fit_t2tx<- survfit(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_ASXL1, data=t2tx)
my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsurv <- ggsurvplot(
  fit_t2tx,
  data = t2tx,
  risk.table = TRUE, fontsize = 6.5,
  pval = FALSE,
  pval.coord = c(0.1, 0.04), #P-value position
  conf.int = FALSE,
  palette = c("#01436D", "#AE1D1E" ), #blue MRD-negative, red MRD-positive
  xlim = c(0,10), # present narrower X axis, but not affect survival estimates.
  xlab = "Time from alloHCT (years)",
  ylab = "Relapse-free survival probability",
  break.time.by = 1, 
  ggtheme = my_theme,
  risk.table.y.text.col = T,
  risk.table.height = 0.225,
  risk.table.y.text = FALSE, 
  ncensor.plot = FALSE, # plot the number of censored subjects at time t
  legend.labs = c("MRD-negative", "MRD-positive"),
  legend.title = element_blank()
)

#legend
ggsurv$plot <- ggsurv$plot +
  theme(
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)), 
    axis.text.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, margin = margin(r = 15)),
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.direction = "horizontal",
    legend.position = c(0.5, 0.1),
    legend.justification = c(0.5, 1),
    #legend.background = element_blank(), #no white frame
    legend.background = element_rect(fill = "white", color = NA), #graph was crossing legend
    legend.margin = margin(t = 2, r = 5, b = 2, l = 5), 
    legend.key = element_rect(fill = "white")
  )

ggsurv$table <- ggsurv$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain") 
  )
##deactivate part until ggsurv to avoid saving as PDF ##
pdf(output_path, width = 6, height = 6) # using inches
print(ggsurv, newpage = FALSE)
dev.off()
ggsurv
```
Calculate the p value with logrank test and Cox regression.
```{r}
survdiff(Surv(RFS_fromTx_years_censored, RFS_status)~t2tx$t2_MRD_ASXL1, data=t2tx)
coxph(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_ASXL1, data=t2tx) %>% 
  tbl_regression(exp = TRUE)
```
CIR t2tx (calculated from Tx)
```{r}
#SAVING AS PDF, DEACTIVE BY USING #
output_path_cir <- "//Users/maximilianbader/Library/Mobile Documents/com~apple~CloudDocs/Universitaet/AG Heuser/analysis/2025/paper/figures/survival analyses/figures MH v2/251206_t2tx_ASXL1only_CIR_t2tx_v1.pdf"
dir.create(dirname(output_path_cir), showWarnings = FALSE, recursive = TRUE)

my_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ci.all <- cuminc(ftime = t2tx$RFS_fromTx_years_censored, fstatus = t2tx$CIR_event, group = t2tx$t2_MRD_ASXL1)

fig <- ggcompetingrisks(
  fit = ci.all,
  multiple_panels = FALSE,
  conf.int = FALSE,
  pval = TRUE,
  xlim = c(0, 10),
  ylim = c(0, 1),
  xlab = "Time from alloHCT (years)",
  ylab = "Cumulative Incidence"
)

fig$data$combined_label <- paste(fig$data$group, fig$data$event, sep = "_")

draw_order <- c("1_1", "1_2", "0_2", "0_1")
fig$data$combined_label <- factor(fig$data$combined_label, levels = draw_order)
fig$data <- fig$data[order(match(fig$data$combined_label, draw_order)), ]

fig$mapping <- aes(x = time, y = est, color = combined_label, linetype = combined_label)
fig$layers <- NULL 

fig <- fig +
  geom_step(linewidth = 1.5) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_manual(
    name = "",
    breaks = c("1_1", "0_1", "1_2", "0_2"),
    labels = c(
      "0_1" = "CIR MRD-neg",
      "0_2" = "NRM MRD-neg",
      "1_1" = "CIR MRD-pos",
      "1_2" = "NRM MRD-pos"
    ),
    values = c(
      "0_1" = "#01436D", 
      "0_2" = "#e18727",
      "1_1" = "#AE1D1E",
      "1_2" = "#20854e"
    ),
    guide = guide_legend(nrow = 2, override.aes = list(linetype = "solid"))
  ) +
  scale_linetype_manual(
    values = c(
      "0_1" = "dashed",
      "0_2" = "solid",
      "1_1" = "solid",
      "1_2" = "solid"
    ),
    guide = "none"
  )

fit_rfs <- survfit(Surv(RFS_fromTx_years_censored, RFS_status) ~ t2tx$t2_MRD_ASXL1, data = t2tx)

ggsurv_table_obj <- ggsurvplot(
  fit_rfs,
  data = t2tx,
  risk.table = TRUE,
  fontsize = 6.5,
  risk.table.height = 0.225,
  palette = c("#01436D", "#AE1D1E"),
  xlim = c(0, 10),
  break.time.by = 1,
  ggtheme = my_theme,
  risk.table.y.text = FALSE,
  risk.table.y.text.col = T,
  legend.labs = c("MRD-negative", "MRD-positive")
)

final_plot <- ggsurv_table_obj
final_plot$plot <- fig

final_plot$plot <- final_plot$plot +
  my_theme +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 18, color = "black", face = "plain"),
    axis.title.x = element_text(size = 18, margin = margin(t = 5)),
    axis.text.x = element_text(size = 18, color = "black"), 
    axis.title.y = element_text(size = 18, margin = margin(r = 15)), 
    axis.text.y = element_text(size = 18, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.title = element_blank(),
    legend.position = c(0.5, 1.005), 
    legend.justification = c(0.5, 1),
    legend.background = element_blank()
  )

final_plot$table <- final_plot$table +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 18, color = "black", face = "plain"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 11.4, unit = "pt"),
    legend.text = element_text(size = 18, face = "plain")
  )

pdf(output_path_cir, width = 6, height = 6)
print(final_plot, newpage = FALSE)
dev.off()
print(final_plot)

fine_gray_test_cir <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 1, t2tx$t2_MRD_ASXL1)
summary(fine_gray_test_cir)

fine_gray_test_nrm <- crr(t2tx$RFS_fromTx_years_censored, t2tx$CIR_event == 2, t2tx$t2_MRD_ASXL1)
summary(fine_gray_test_nrm)
```
